---
title: ""
author: ""
date: ""
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 3
    smooth_scroll: false
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(plotly)
library(ggplot2)
library(data.table)
library(DESeq2)
library(reactable)
library(hictoolsr)
setwd("~/Documents/MED4_collab")
```

Loops were firstly called using the [ABC Model](https://doi.org/10.1038/s41588-019-0538-0). Since this method integrates ATAC-seq, H3K27ac and gene expression data when calling enhancer-promoter (E-P) pairs, we can be more confident we are truly calling loops. After defining E-P loops, differential loop calling was performed using a modified DESeq2 approach via [Hictoolsr](https://github.com/EricSDavis/hictoolsr).

# ABC model results
```{r,ABC_MODEL,echo=FALSE,fig.cap="The number of significant loops (FDR < 0.05) is reported in each condition."}
xfun::embed_file("~/Documents/MED4_collab/data/ABC_model_out/predictions/Enhancer_promoter_loops.csv")
```

# Differential loop analysis

Differential loop analysis was performed at 1000kb resolution (ie. regions were binned at 10kb windows).

## Raw counts {.tabset}

### Counts
```{r,rawCounts, echo=FALSE}
load("~/Documents/MED4_collab/data/hictoolsr/loopCounts.1000bp.RData")
df <- loopCounts %>% as.data.frame()
df %>% reactable()
```

### E-P pairs
```{r,E_P_PAIRS, echo=FALSE}
bedpe <- fread("~/Documents/MED4_collab/data/ABC_model_out/ABC.enhancer_promoter.bed2d") %>%
  as_ginteractions( keep.extra.columns = TRUE,
  starts.in.df.are.0based = FALSE
)
bedpe %>% as.data.frame() %>% reactable()
```

## DESeq2 res
```{r,DESEQ2_RES, include=FALSE}
df$loops <- stringr::str_c(df$seqnames1,sep = ".",df$start1,df$end1,df$start2,df$end2)
cnts <- df %>%
  distinct() %>%
  dplyr::select(c(20,11:19)) %>%
  column_to_rownames("loops")
colnames(cnts) <- gsub(".hic","",colnames(cnts))
# mdat
mdat <- data.frame(kind=colnames(cnts))
mdat$condition <- "KD"
mdat[7:9,"condition"] <- "NTC"
mdat <- mdat %>%
  column_to_rownames("kind")
## Build DESeq data set
for (n in 1:9) {
  cnts[,n] <- as.integer(cnts[,n])
}
dds <- DESeqDataSetFromMatrix(countData = cnts,
                              colData = mdat,
                              design = ~ condition)
dds <- DESeq(dds)
vst <- varianceStabilizingTransformation(dds)
```

```{r,FINAL_RES, echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
res <- results(dds,contrast=c("condition","KD","NTC")) %>%
  as.data.frame() %>%
  rownames_to_column("loops")
res %>% reactable()
```

## PCA {.tabset}

### Top 500 most variable loops
```{r,PCA_500, echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rv <- rowVars(assay(vst))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
pc <- prcomp(t(assay(vst)[select,]))
condition <- mdat$condition
scores <- data.frame(pc$x, condition)
percentage <- round(pc$sdev / sum(pc$sdev) * 100, 2)
percentage <- paste( colnames(scores), "(", paste( as.character(percentage), "%", ")", sep="") )
p <- plot_ly(scores,x=scores$PC1,y=scores$PC2,text=rownames(scores),mode="markers",color=factor(condition),marker=list(size=11),colors = "Dark2")
p <- layout(p,title="",   xaxis = list(title = percentage[1]),
            yaxis = list(title = percentage[2]))
p
```

### All loops
```{r,PCA_ALL, echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rv <- rowVars(assay(vst))
pc <- prcomp(t(assay(vst)))
condition <- mdat$condition
scores <- data.frame(pc$x, condition)
percentage <- round(pc$sdev / sum(pc$sdev) * 100, 2)
percentage <- paste( colnames(scores), "(", paste( as.character(percentage), "%", ")", sep="") )
p <- plot_ly(scores,x=scores$PC1,y=scores$PC2,text=rownames(scores),mode="markers",color=factor(condition),marker=list(size=11),colors = "Dark2")
p <- layout(p,title="",   xaxis = list(title = percentage[1]),
            yaxis = list(title = percentage[2]))
p
```

## MA plot

```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
DESeq2::plotMA(dds,colSig="blue",xlab="Mean of Normalized Counts",alpha=0.05,ylim=c(-6,6))
```

