---
title: ""
author: ""
date: ""
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 3
    smooth_scroll: false
    theme: cerulean
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F,
  message = F,
  warning = F,
  out.width = '100%'
)
library(data.table)
library(tidyverse)
library(highcharter)
library(plotly)
library(smacof)
library(uwot)
library(pals)
library(heatmaply)
library(ggdist)
library(rtracklayer)
library(crosstalk)
library(trelliscopejs)
library(DT)
library(reticulate)
np <- import('numpy')

plot_ly()
highchart()
heatmaply(matrix(1:4, 2))
```

# Mapping {.tabset .tabset-pills}

## All {.tabset}

```{r}
d <- list.files('~/Documents/MED4_collab/data/qc/stats', full.names = T) %>% 
  setNames(., sub('.txt', '', basename(.))) %>%
  lapply(fread) %>% 
  bind_rows(.id = 'samp') %>%
  dplyr::filter(!grepl('dist_freq|chrom_freq|pair_types', V1)) %>%
  mutate(V2 = as.integer(V2), 
         V1 = sub('summary/', '', V1)) %>%
  dplyr::filter(!grepl('^frac_', V1)) %>% 
  pivot_wider(names_from = V1, values_from = V2) %>%
  mutate(type = sub('-.*', '', samp))

pd <- d[,c('samp','type','total_unmapped','total_single_sided_mapped','total_mapped')] %>%
  pivot_longer(-c(samp,type), values_to = 'num') %>%
  mutate(name = sub('total_', '', name)) %>%
  group_by(samp) %>%
  mutate(perc = 100 * num / sum(num),
         name = factor(name, c('unmapped','single_sided_mapped','mapped'))) %>%
  ungroup()


```


### Percentage

```{r}
  
hchart(pd, 'column', hcaes(x = samp, y = perc, group = name),
       stacking = "normal") %>%
  hc_xAxis(title = list(text = 'Sample')) %>%
  hc_yAxis(title = list(text = 'Percentage of pairs'),
           min = 0, max = 100) %>%
  hc_chart(zoomType = 'xy') %>%
  hc_exporting(
        enabled = TRUE,
        buttons = list(
          customButton = list(text = 'Linear',onclick = JS("function() {this.yAxis[0].update({type: 'linear'});}")),
          customButton2 = list(text = 'Log', onclick = JS("function() {this.yAxis[0].update({type: 'logarithmic'});}"))
        )
      )

```


### Count

```{r}
hchart(pd, 'column', hcaes(x = samp, y = num, group = name),
       stacking = "normal") %>%
  hc_xAxis(title = list(text = 'Sample')) %>%
  hc_yAxis(title = list(text = 'Number of pairs')) %>%
  hc_chart(zoomType = 'xy') %>%
  hc_exporting(
        enabled = TRUE,
        buttons = list(
          customButton = list(text = 'Linear',onclick = JS("function() {this.yAxis[0].update({type: 'linear'});}")),
          customButton2 = list(text = 'Log', onclick = JS("function() {this.yAxis[0].update({type: 'logarithmic'});}"))
        )
      )
```


## Mapped {.tabset}

```{r}

pd <- d[,c('samp','type','total_dups','total_nodups')] %>%
  pivot_longer(-c(samp,type), values_to = 'num') %>%
  mutate(name = sub('total_', '', name)) %>%
  group_by(samp) %>%
  mutate(perc = 100 * num / sum(num)) %>%
  ungroup()


```

### Percentage

```{r}
  
hchart(pd, 'column', hcaes(x = samp, y = perc, group = name),
       stacking = "normal") %>%
  hc_xAxis(title = list(text = 'Sample')) %>%
  hc_yAxis(title = list(text = 'Percentage of pairs'),
           min = 0, max = 100) %>%
  hc_chart(zoomType = 'xy') %>%
  hc_exporting(
        enabled = TRUE,
        buttons = list(
          customButton = list(text = 'Linear',onclick = JS("function() {this.yAxis[0].update({type: 'linear'});}")),
          customButton2 = list(text = 'Log', onclick = JS("function() {this.yAxis[0].update({type: 'logarithmic'});}"))
        )
      )

```


### Count

```{r}
hchart(pd, 'column', hcaes(x = samp, y = num, group = name),
       stacking = "normal") %>%
  hc_xAxis(title = list(text = 'Sample')) %>%
  hc_yAxis(title = list(text = 'Number of pairs')) %>%
  hc_chart(zoomType = 'xy') %>%
  hc_exporting(
        enabled = TRUE,
        buttons = list(
          customButton = list(text = 'Linear',onclick = JS("function() {this.yAxis[0].update({type: 'linear'});}")),
          customButton2 = list(text = 'Log', onclick = JS("function() {this.yAxis[0].update({type: 'logarithmic'});}"))
        )
      )
```


## Unique {.tabset}


```{r}

pd <- d[,c('samp','type','cis','trans')] %>%
  pivot_longer(-c(samp,type), values_to = 'num') %>%
  mutate(name = sub('total_', '', name)) %>%
  group_by(samp) %>%
  mutate(perc = 100 * num / sum(num),
         name = factor(name, c('trans','cis'))) %>%
  ungroup()


```

### Percentage

```{r}
  
hchart(pd, 'column', hcaes(x = samp, y = perc, group = name),
       stacking = "normal") %>%
  hc_xAxis(title = list(text = 'Sample')) %>%
  hc_yAxis(title = list(text = 'Percentage of pairs'),
           min = 0, max = 100) %>%
  hc_chart(zoomType = 'xy') %>%
  hc_exporting(
        enabled = TRUE,
        buttons = list(
          customButton = list(text = 'Linear',onclick = JS("function() {this.yAxis[0].update({type: 'linear'});}")),
          customButton2 = list(text = 'Log', onclick = JS("function() {this.yAxis[0].update({type: 'logarithmic'});}"))
        )
      )

```


### Count

```{r}
hchart(pd, 'column', hcaes(x = samp, y = num, group = name),
       stacking = "normal") %>%
  hc_xAxis(title = list(text = 'Sample')) %>%
  hc_yAxis(title = list(text = 'Number of pairs')) %>%
  hc_chart(zoomType = 'xy') %>%
  hc_exporting(
        enabled = TRUE,
        buttons = list(
          customButton = list(text = 'Linear',onclick = JS("function() {this.yAxis[0].update({type: 'linear'});}")),
          customButton2 = list(text = 'Log', onclick = JS("function() {this.yAxis[0].update({type: 'logarithmic'});}"))
        )
      )
```


## Complexity

```{r}
hchart(d, 'column', hcaes(x = samp, y = complexity_naive),
       stacking = "normal") %>%
  hc_xAxis(title = list(text = 'Sample')) %>%
  hc_yAxis(title = list(text = 'Estimated complexity')) %>%
  hc_chart(zoomType = 'xy') %>%
  hc_exporting(
        enabled = TRUE,
        buttons = list(
          customButton = list(text = 'Linear',onclick = JS("function() {this.yAxis[0].update({type: 'linear'});}")),
          customButton2 = list(text = 'Log', onclick = JS("function() {this.yAxis[0].update({type: 'logarithmic'});}"))
        )
      )
```

# {.unlisted .unnumbered .toc-ignore}

---

# Matrix similarity

## Matrix

```{r}

clrs <- tableau20()[rep(c(1,3,5), each = 3)]


ps <- list.files('~/Documents/MED4_collab/data/qc/exp', pattern = '1000.cis.(der|log).tsv', 
           recursive = T, full.names = T) %>%
  split(., sub('.*\\.(.*)\\.tsv', '\\1', .)) %>%
  lapply(function(x) {
    split(x, basename(dirname(x))) %>%
      lapply(function(y) {
        fread(y) -> dd
        dd[,!(colnames(dd) %in% c('diag_bin_start', 'diag_bin_end')),with=F]
      })
  })

pd <- ps$der %>%
  rbindlist(idcol = 'samp') %>%
  mutate(type = sub('-.*', '', samp) %>% factor(c('NTC','M1','M4'))) %>%
  arrange(type, samp) %>%
  dplyr::filter(type != samp) %>%
  mutate(rep = sub('.*-', 'Rep ', samp),
         samp = fct_inorder(samp)) %>%
  dplyr::filter(s_bp < 1.5e8) 

ann <- pd[,c('type', 'samp')] %>%
  distinct(samp, type) %>%
  column_to_rownames('samp')
clrss <- ann %>% mutate(clr = clrs) %>% distinct() %>% deframe()
md <- distinct(pd[,c('samp','type','rep')])

f <- fread('~/Documents/MED4_collab/ref/mm10.chrom.sizes')$V2 %>% {./sum(.)}

hc <- list.files('~/Documents/MED4_collab/data/hicrep/50000/5', full.names = T) %>%
  tibble(f = .) %>%
  group_by(f) %>%
  do(., tibble(v = sum(fread(.$f, skip = 2)[[1]] * f))) %>%
  ungroup() %>%
  mutate(f = sub('.txt', '', basename(f))) %>%
  separate(f, c('s1','s2'), '\\.') %>%
  {rbind(., .[,c(2,1,3)] %>% `names<-`(c('s1','s2','v')))} %>%
  {rbind(., tibble(s1 = unique(.$s1, .$s2), s2 = s1, v = 1))} %>%
  distinct() %>%
  arrange(s1, s2) %>%
  pivot_wider(names_from = 's2', values_from = 'v') %>%
  column_to_rownames('s1') %>%
  .[levels(pd$samp), levels(pd$samp)]

hc %>%
  heatmaply(row_side_colors = ann,
            row_side_palette = clrss,
            col_side_colors = ann,
            col_side_palette = clrss)

```

## UMAP

```{r}
as.dist(1 - hc) %>%
  umap(n_components = 3, n_neighbors = 5) %>%
  as.data.frame() %>%
  mutate(samp = colnames(hc)) %>%
  merge(md) %>%
  plot_ly(x = ~V1, y = ~V2, z = ~V3, type = 'scatter3d', mode = 'markers',
          color = ~type, colors = clrss) %>%
  layout(scene = list(xaxis = list(title = 'Dim 1'),
                      yaxis = list(title = 'Dim 2'),
                      zaxis = list(title = 'Dim 3')))
```

# {.unlisted .unnumbered .toc-ignore}

---

# Contact decay {.tabset}

## P(s)

```{r}
ps$log %>%
  rbindlist(idcol = 'samp') %>%
  merge(md) %>%
  dplyr::filter(s_bp < 1.5e8) %>%
  hchart('line', hcaes(x = s_bp, y = balanced.avg, group = samp)) %>%
  hc_xAxis(type = 'logarithmic', title = list(text = 'Genomic separation')) %>%
  hc_yAxis(type = 'logarithmic', title = list(text = 'Contact probability')) %>%
  hc_colors(clrs) %>%
  hc_chart(zoomType = 'xy') %>%
  hc_plotOptions(line = list(marker = list(enabled = F)))
```

## P'(s)

```{r}
hchart(pd, 'line', hcaes(x = s_bp, y = slope, group = samp)) %>%
  hc_xAxis(type = 'logarithmic', title = list(text = 'Genomic separation')) %>%
  hc_yAxis(title = list(text = 'Contact probability decay')) %>%
  hc_colors(clrs) %>%
  hc_chart(zoomType = 'xy') %>%
  hc_plotOptions(line = list(marker = list(enabled = F)))
```

# {.unlisted .unnumbered .toc-ignore}

---

# Compartments

## Global similarity {.tabset .tabset-pills}

### Correlation matrix

```{r}
bsz <- '50000'
m <- list.files('~/Documents/MED4_collab/data/qc/eig', pattern = paste0('^', bsz, '.cis.vecs.tsv'), 
                full.names = T, recursive = T) %>%
  setNames(., basename(dirname(.))) %>%
  lapply(function(x) fread(x)$E1) %>%
  bind_cols() %>%
  .[,levels(pd$samp)] %>%
  na.omit()


cor(m) %>%
  heatmaply(row_side_colors = ann,
            row_side_palette = clrss,
            col_side_colors = ann,
            col_side_palette = clrss)
```

### PCA

```{r}
res <- t(m) %>% 
  prcomp()

axs <- summary(res)$importance[2,] %>% 
  mapply(function(x, y) {sprintf('%s (%.2f%%)', y, x * 100)}, ., names(.))

res$x %>%
  as.data.frame() %>%
  rownames_to_column('samp') %>%
  merge(md) %>%
  plot_ly(x = ~PC1, y = ~PC2, z = ~PC3, type = 'scatter3d', mode = 'markers',
          color = ~type, colors = clrss) %>%
  layout(scene = list(xaxis = list(title = axs[1]),
                      yaxis = list(title = axs[2]),
                      zaxis = list(title = axs[3])))
# pd <- res$x %>%
#   as.data.frame() %>%
#   rownames_to_column('samp') %>%
#   merge(md)
# pd %>%
#   ggplot(aes(x = PC1, y = PC2, color = type)) +
#   geom_point(size = 3, show.legend = T) +
#   geom_text_repel(aes(label=samp),show.legend = FALSE) +
#   geom_hline(yintercept = 0) +
#   geom_vline(xintercept = 0) +
#   xlab(label = "PC1 (45.85%)") +
#   ylab(labe = "PC2 (20.94%)") +
#   labs(color = "Condition") +
#   theme_bw()
# ggsave(dpi = 600,filename = "~/Documents/MED4_collab/data/images/compartment_score.png",units = "in",width = 7,height=4)
```

### UMAP {.tabset}

#### Euclidean

```{r}
umap(t(m), n_components = 3, metric = 'euclidean', n_neighbors = 5) %>%
  as.data.frame() %>%
  mutate(samp = colnames(m)) %>%
  merge(md) %>%
  plot_ly(x = ~V1, y = ~V2, z = ~V3, type = 'scatter3d', mode = 'markers',
          color = ~type, colors = clrss) %>%
  layout(scene = list(xaxis = list(title = 'Dim 1'),
                      yaxis = list(title = 'Dim 2'),
                      zaxis = list(title = 'Dim 3')))
```

#### Correlation

```{r}
umap(t(m), n_components = 3, metric = 'correlation', n_neighbors = 5) %>%
  as.data.frame() %>%
  mutate(samp = colnames(m)) %>%
  merge(md) %>%
  plot_ly(x = ~V1, y = ~V2, z = ~V3, type = 'scatter3d', mode = 'markers',
          color = ~type, colors = clrss) %>%
  layout(scene = list(xaxis = list(title = 'Dim 1'),
                      yaxis = list(title = 'Dim 2'),
                      zaxis = list(title = 'Dim 3')))
```

## {.unlisted .unnumbered .toc-ignore}

---

## Distribution



```{r, dev = 'svg'}
m %>% 
  mutate(idx = 1:n()) %>% 
  pivot_longer(-idx, names_to = 'samp', values_to = 'v') %>%
  mutate(samp = factor(samp, levels(pd$samp))) %>%
  merge(md) %>%
  ggplot(aes(x = samp,y  = v, color = type, fill = type)) +
  geom_hline(yintercept = 0) +
  geom_violin(color = NA, alpha = .5) +
  stat_pointinterval() +
  scale_fill_manual(values = clrss) +
  scale_color_manual(values = clrss) +
  facet_wrap(~type, scales = 'free_x',nrow=1) +
  scale_x_discrete(labels = function(x) sub('Rep', 'Rep ', sub('.*\\ ', '', x))) +
  labs(y = 'Compartment score') +
  coord_cartesian(ylim = c(-2, 2)) +
  scale_y_continuous(breaks = -2:2) +
  theme(plot.background = element_blank(),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        legend.position = 'none',
        axis.line.x = element_line(color = 'black'),
        panel.grid.major.y = element_line(color = 'grey75', linetype = 'dashed'),
        strip.background = element_rect(fill = 'black'),
        strip.text = element_text(color = 'white'),
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text = element_text(color = 'black'))
```




---

## Compartment size

```{r}
o <- list.files('~/Documents/MED4_collab/data/qc/eig', pattern = paste0('^', bsz, '.cis.vecs.tsv'), 
                full.names = T, recursive = T) %>%
  setNames(., basename(dirname(.))) %>%
  lapply(function(x) fread(x)$E1) %>%
  bind_cols() %>%
  cbind(fread(list.files('~/Documents/MED4_collab/data/qc/eig', pattern = paste0('^', bsz, '.cis.vecs.tsv'), 
                full.names = T, recursive = T)[1] , select = 1:3), .) %>%
  pivot_longer(-c('chrom','start','end'), 
               names_to = 'samp', values_to = 'score') %>% 
  dplyr::rename(chr = chrom) %>% 
  filter(grepl('-[123]$', samp)) %>% 
  na.omit() %>% 
  add_count(chr, start) %>% 
  filter(n == max(n)) %>% 
  dplyr::select(-n) 


dm <- o %>%
  mutate(idx = as.integer(start / as.integer(bsz))) %>%
  arrange(samp, chr, idx) 

dl <- distinct(dm, chr, idx) %>%
  split(., .$chr) %>%
  lapply(function(x) {
    group_by(x, gw = cumsum(c(1, diff(idx) != 1))) %>%
      summarise(n = n(), i = idx[1], j = idx[n]) %>%
      ungroup()
  }) %>%
  bind_rows(.id = 'chr') %>%
  group_by(chr) %>%
  top_n(1, n) %>%
  ungroup() %>%
  mutate(l = n * as.integer(bsz) / 1e6)

ac <- split(dm, dm$samp) %>%
  lapply(function(x) {
    split(x, x$chr) %>%
      lapply(function(y) {
        y %>%
          dplyr::filter(y$idx > dl[dl$chr == y$chr[1],]$i &
                          y$idx < dl[dl$chr == y$chr[1],]$j) %>%
          pull(score) %>%
          acf(na.action = na.pass, plot = F,
              lag.max = dl[dl$chr == y$chr[1],]$n) %>%
          {tibble(lag = .$lag[,,1], c = .$acf[,,1])} %>%
          dplyr::filter(lag >= 0) %>%
          mutate(lag = lag * as.integer(bsz),
                 c = abs(c))
      }) %>%
      bind_rows(.id = 'chr') %>%
      dplyr::filter(!(chr %in% c('chrY'))) %>%
      group_by(lag) %>%
      summarise(s = sd(c, na.rm = T) / sqrt(20),
                c = mean(c, na.rm = T),
                lo = c - s, hi = c + s) 
  }) %>%
  bind_rows(.id = 'samp') %>%
  na.omit() %>%
  dplyr::filter(lag != 0)  %>%
  merge(md) %>%
  mutate(samp = factor(samp, levels(md$samp))) %>%
  arrange(samp)

ac %>% 
  filter(between(lag, 0, 3e6)) %>%
  hchart('line', hcaes(x = lag, y = c, group = samp)) %>%
  hc_colors(clrs) %>%
  hc_chart(zoomType = 'xy') %>%
  hc_xAxis(title = list(text = 'Lag')) %>%
  hc_yAxis(title = list(text = 'Autocorrelation')) %>%
  hc_plotOptions(line = list(marker = list(enabled = F)))
```

---

# Insulation

## Correlation

```{r}
i <- list.files('~/Documents/MED4_collab/data/qc/ins', full.names = T, pattern = '25000.1000000.tsv', recursive = T) %>%
  setNames(., basename(dirname(.))) %>%
  lapply(fread)

lapply(i, `[[`, 5) %>%
  .[levels(pd$samp)] %>%
  bind_cols() %>%
  na.omit() %>%
  cor() %>%
  heatmaply(row_side_colors = ann,
            row_side_palette = clrss,
            col_side_colors = ann,
            col_side_palette = clrss)

```

---

## # of insulating loci

```{r}
lapply(i, function(x) {tibble(n = sum(na.omit(x[[7]]) > .1))}) %>% 
  bind_rows(.id = 'samp') %>%
  merge(md) %>%
  mutate(samp = factor(samp, levels(md$samp))) %>%
  arrange(samp) %>%
  mutate(clr = clrs) %>%
  hchart('point', hcaes(x = samp, y = n, color = clr)) %>%
  hc_xAxis(title = NULL) %>%
  hc_yAxis(title = list(text = '# of bins with boundary strength >0.1 <br> @25kb resolution with 1mb window size',
                        useHTML = T), useHTML = T) %>%
  hc_tooltip(headerFormat = '<span style="font-size: 10px">{point.key}</span><br/>')
```

---




