---
title: ""
author: ""
date: ""
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 3
    smooth_scroll: false
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(ggpubr)
library(gridExtra)
library(grid)
library(cowplot)
library(reactable)
library(data.table)
library(reticulate)
library(sciplot)
library(plotly)
library(matrixStats)
library(SummarizedExperiment)
library(RColorBrewer)
library(heatmaply)
library(ggrepel)
library(tidyverse)
library(gprofiler2)
library(patchwork)
library(DESeq2)
library(apeglm)
library(fgsea)
library(msigdbr)
library(compiler)
library(shiny)
library(tippy)
library(msigdbr)
# library(GSVA)
library(pals)
library(ggtext)
library(broom)
library(DT)
library(EnsDb.Mmusculus.v79)
library(org.Mm.eg.db)
library(eulerr)
library(GeneOverlap)
setwd("~/Documents/MED4_collab")
```

# PCA
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
load("~/Documents/MED4_collab/data/RData/RNAseq/vst_condition.RData")
load("~/Documents/MED4_collab/data/RData/RNAseq/mdat.RData")
rv <- rowVars(assay(vst))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
pc <- prcomp(t(assay(vst)[select,]))
condition <- mdat$condition
scores <- data.frame(pc$x, condition)
percentage <- round(pc$sdev / sum(pc$sdev) * 100, 2)
percentage <- paste( colnames(scores), "(", paste( as.character(percentage), "%", ")", sep="") )
p <- plot_ly(scores,x=scores$PC1,y=scores$PC2,text=rownames(scores),mode="markers",color=factor(condition),marker=list(size=11),colors = "Dark2")
p <- layout(p,title="",   xaxis = list(title = percentage[1]),
            yaxis = list(title = percentage[2]))
p
```

# Hierarchical clustering
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rv <- rowVars(assay(vst))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
gplots::heatmap.2(assay(vst)[select,], margins = c(10, 6),trace = "none", col = viridis(100),srtCol = 70,cexCol = 0.8)
#heatmaply(assay(rld)[select,],seriate = "mean",row_dend_left=TRUE,plot_method="plotly",hide_colorbar = TRUE)
```

# Correlation Heatmap
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rv <- rowVars(assay(vst))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
#sampleDists <- stats::dist(t(assay(rld[select])),method = "euclidean")
#sampleDistMatrix <- as.matrix(sampleDists)
#colors <- colorRampPalette( rev(brewer.pal(9, "YlOrRd")) )(255)
#heatmaply(sampleDistMatrix,
        #  clustering_distance_rows=sampleDists,
       #   clustering_distance_cols=sampleDists,
       #   col=colors,showticklabels = FALSE,key.title = "Euclidean distance",row_dend_left = TRUE,dend_hoverinfo #= FALSE)

heatmaply_cor(
  cor(assay(vst[select,])),showticklabels = TRUE,key.title = "",row_dend_left = FALSE)
```

# Volcano {.tabset}
```{r volcano_plot, include=FALSE}
shrunken.logfc <- function(dds,group,sampA,ref_sampB){ #group is similar to condition
  df_shrunken <- lfcShrink(dds,contrast=c(group,sampA,ref_sampB),type="ashr")
  df <- df_shrunken %>%
    data.frame() %>%
    na.omit() %>%
    rownames_to_column('ensembl_gene_id')
  geneSymbols <- ensembldb::select(EnsDb.Mmusculus.v79, keys=df$ensembl_gene_id, keytype = "GENEID", columns = c("SYMBOL","GENEID"))
  colnames(geneSymbols) <- c("gene","ensembl_gene_id")
  rd <- full_join(df,geneSymbols,by="ensembl_gene_id")
  return(rd)
}
#
volc <- function(r, x, y, ylab, ttl) {
  d <- as.data.frame(r) %>%
    dplyr::rename(x = !!x, y = !!y) %>%
    mutate(kind = case_when((abs(x) > 2 & y < .05) ~ 'DE',
                            abs(x) > 2 ~ '|logFC|>2',
                            y < .05 ~ 'FDR<0.05',
                            T ~ 'NS'),
           y = -log10(y))
  ct <- d %>% 
    dplyr::filter(kind == 'DE') %>%
    mutate(up = x > 0) %>%
    dplyr::count(up) %>%
    mutate(x = ifelse(up, Inf, -Inf),
           y = Inf,
           h = as.numeric(up))
  ggplot(d, aes(x, y, color = kind)) +
    geom_vline(xintercept = c(-2, 2), linetype = 'dashed') +
    geom_hline(yintercept = -log10(0.05), linetype = 'dashed') +
    geom_point(alpha = .5) +
    geom_label(aes(x = x, y = y, label = n, hjust = h),
              vjust = 1, data = ct, inherit.aes = F) +
     geom_text_repel(aes(label = gene), data = subset(d, kind == "DE"),max.overlaps = 10,show.legend = F, min.segment.length = 0) + scale_y_continuous() +
    scale_color_manual(values = c('orange','forestgreen', 'red',  'black')) +
    labs(x = x, y = ylab, title = ttl) + theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.title = element_blank())
}
```

## M1 vs NTC
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
load("~/Documents/MED4_collab/data/RData/RNAseq/dds_condition.RData")
load("~/Documents/MED4_collab/data/RData/RNAseq/rd_M1_v_NTC.RData")
# cdat <- cdat %>%
#   rownames_to_column("ensembl_gene_id")
#
srd <- shrunken.logfc(dds = dds_condition,group = "condition","M1","NTC")
volc(srd,'log2FoldChange','padj','-log10(FDR)','')
```

## M4 vs NTC
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
load("~/Documents/MED4_collab/data/RData/RNAseq/rd_M4_v_NTC.RData")
# cdat <- cdat %>%
#   rownames_to_column("ensembl_gene_id")
#
srd <- shrunken.logfc(dds = dds_condition,group = "condition","M4","NTC")
volc(srd,'log2FoldChange','padj','-log10(FDR)','')
```

## KD vs NTC
M1 and M4 were pooled together and compared against NTC. *Note the unbalanced comparison: n=6 for MED4KD (3 reps for M1 and 3 reps for M4) vs. n=3 for NTC

```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
load("~/Documents/MED4_collab/data/RData/RNAseq/rd_KD_v_NTC.RData")
load("~/Documents/MED4_collab/data/RData/RNAseq/dds_treatment.RData")
# cdat <- cdat %>%
#   rownames_to_column("ensembl_gene_id")
#
srd <- shrunken.logfc(dds = dds_treatment,group = "treatment","KD","Control")
volc(srd,'log2FoldChange','padj','-log10(FDR)','')
```

# Pathway Analysis {.tabset}

Using criteria of log2FC > 2 and FDR < 0.05
```{r include=FALSE}
pathway_analysis <- function(x,fc_cutoff){ #x is dge df OR rd
  up <- gost(x$ensembl_gene_id[x$padj < .05 & x$log2FoldChange > fc_cutoff],
             organism = 'mmusculus',
             custom_bg = x$ensembl_gene_id,
             user_threshold = 0.05,
             significant = TRUE,
             evcodes = TRUE,
             exclude_iea = TRUE,
             correction_method = "fdr",
             sources = c("GO","KEGG","REAC","TF","CORUM"))
  down <- gost(x$ensembl_gene_id[x$padj < .05 & x$log2FoldChange < -fc_cutoff],
             organism = 'mmusculus',
             custom_bg = x$ensembl_gene_id,
             user_threshold = 0.05,
             significant = TRUE,
             evcodes = TRUE,
             exclude_iea = TRUE,
             correction_method = "fdr",
             sources = c("GO","KEGG","REAC","TF","CORUM"))
  up_reactable <- up$result %>% dplyr::select(source, term_name, term_id, intersection_size, p_value)
  down_reactable <- down$result %>% dplyr::select(source, term_name, term_id, intersection_size, p_value)
  return(list(up=up,down=down,up_reactable=up_reactable,down_reactable=down_reactable))
}
```

## M1 v NTC {.tabset}

### Upregulated
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
# load("~/Documents/MED4_collab/data/RData/RNAseq/pa_M1_v_NTC.RData")
pa_M1_v_NTC <- pathway_analysis(x = rd_M1_v_NTC,fc_cutoff = 2)
gostplot(pa_M1_v_NTC$up,capped = FALSE,interactive = TRUE)
reactable(pa_M1_v_NTC$up_reactable)
```

### Downregulated
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
gostplot(pa_M1_v_NTC$down,capped = FALSE,interactive = TRUE)
reactable(pa_M1_v_NTC$down_reactable)
```

## M4 v NTC {.tabset}

### Upregulated
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
# load("~/Documents/MED4_collab/data/RData/RNAseq/pa_M4_v_NTC.RData")
pa_M4_v_NTC <- pathway_analysis(rd_M4_v_NTC,fc_cutoff = 2)
gostplot(pa_M4_v_NTC$up,capped = FALSE,interactive = TRUE)
reactable(pa_M4_v_NTC$up_reactable)
```

### Downregulated
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
gostplot(pa_M4_v_NTC$down,capped = FALSE,interactive = TRUE)
reactable(pa_M4_v_NTC$down_reactable)
```

## KD v NTC {.tabset}

### Upregulated
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
pa_KD_v_NTC <- pathway_analysis(rd_KD_v_NTC,fc_cutoff = 2)
gostplot(pa_KD_v_NTC$up,capped = FALSE,interactive = TRUE)
reactable(pa_KD_v_NTC$up_reactable)
```

### Downregulated
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
gostplot(pa_KD_v_NTC$down,capped = FALSE,interactive = TRUE)
reactable(pa_KD_v_NTC$down_reactable)
```

# GSEA {.tabset}
```{r include=FALSE}
#gsea
gsea <- function(x){
  e2s <- AnnotationDbi::select(org.Mm.eg.db,
                             key = x$ensembl_gene_id,
                             columns = "SYMBOL",
                             keytype = "ENSEMBL") %>%
  na.omit() %>%
  deframe()
  # use DE test statistic to rank genes
  s <- x %>% 
  mutate(symb = e2s[ensembl_gene_id]) %>%
  na.omit() %>% 
  group_by(symb) %>%
  summarise(stat = mean(stat)) %>%
  deframe()
  # pathways
  p <- msigdbr(species = "Mus musculus", category = "H") %>%
  mutate(gs_name = sub('^HALLMARK_', '', gs_name)) %>%
    {split(.$gene_symbol, .$gs_name)}
  # GSEA
  r <- fgsea(pathways = p, stats = s, eps = 0.0)
  plot.r <- r %>%
  arrange(NES) %>% 
  mutate(pathway = fct_inorder(pathway)) %>% 
  ggplot(aes(x = pathway, y = NES)) + 
  geom_col(aes(fill = -log10(padj))) +
  scale_fill_viridis_c() +
  coord_flip() +
  labs(x = "Pathway",
       y = "Normalized enrichment score")
  top_hit <- plotEnrichment(p[[arrange(r, padj)$pathway[1]]], s) +
  ggtitle(arrange(r, padj)$pathway[1])
  gsea_table <- arrange(r, padj) %>%
  dplyr::select(pathway, padj, ES, NES, size, leadingEdge) %>%
  mutate_at(c('pathway', 'padj', 'ES', 'NES'), function(x) {
    formatC(x, digits = 3, format = 'g')
  }) %>%
  reactable(
    searchable = T,
    highlight = T,
    wrap = F,
    resizable = T,
    striped = T,
    paginationType = "jump",
    showPageSizeOptions = T,
    defaultPageSize = 10,
    columns = list(
      leadingEdge = colDef(
        html = T,
        cell =  function(value, index, name) {
          value <- paste(value, collapse = ', ')
          div(
            style = "cursor: info;
                     white-space: nowrap;
                     overflow: hidden;
                     text-overflow: ellipsis;",
            tippy(text = value, tooltip = value)
          )
        }
      )
    )
  )
  return(list(s=s,p=p,r=r,plot.r=plot.r,top_hit=top_hit,gsea_table=gsea_table))
}
```

## M1 v NTC 
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
gsea_M1_v_NTC <- gsea(x = rd_M1_v_NTC)
plot(gsea_M1_v_NTC$plot.r)
gsea_M1_v_NTC$gsea_table
```

## M4 v NTC
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
gsea_M4_v_NTC <- gsea(x = rd_M4_v_NTC)
plot(gsea_M4_v_NTC$plot.r)
gsea_M4_v_NTC$gsea_table
```

## KD v NTC
```{r echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
gsea_KD_v_NTC <- gsea(x = rd_KD_v_NTC)
plot(gsea_KD_v_NTC$plot.r)
gsea_KD_v_NTC$gsea_table
```