---
title: ""
author: ""
date: ""
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 3
    smooth_scroll: false
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(HiCDCPlus)
library(ggplot2)
library(knitr)
library(reshape2)
library(tidyverse)
library(rtracklayer)
library(isoband)
library(sf)
library(MASS)
library(lwgeom)
library(ggrepel)
library(hexbin)
library(ggrastr)
library(viridis)
library(pals)
library(patchwork)
library(highcharter)
library(dplyr)
library(GenomicRanges)
library(reactable)
library(data.table)
#
library(DESeq2)
library(edgeR)
#
library(plotly)
#
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
library(org.Mm.eg.db)
library(ChIPseeker)
setwd("~/Documents/MED4_collab")
```

```{r,loading_R_objects,include = FALSE}
load("~/Documents/MED4_collab/data/HiCDCPlus/counts/df_list_V1.RData")
volc <- function(r, x, y, ylab, ttl) {
  d <- as.data.frame(r) %>%
    dplyr::rename(x = !!x, y = !!y) %>%
    mutate(kind = case_when((abs(x) > 1 & y < .05) ~ 'DE',
                            abs(x) > 1 ~ '|logFC|>1',
                            y < .05 ~ 'FDR<0.05',
                            T ~ 'NS'),
           y = -log10(y))
  ct <- d %>% 
    dplyr::filter(kind == 'DE') %>%
    mutate(up = x > 0) %>%
    dplyr::count(up) %>%
    mutate(x = ifelse(up, Inf, -Inf),
           y = Inf,
           h = as.numeric(up))
  ggplot(d, aes(x, y, color = kind)) +
    geom_vline(xintercept = c(-1, 1), linetype = 'dashed') +
    geom_hline(yintercept = -log10(0.05), linetype = 'dashed') +
    geom_point(alpha = .5) +
    geom_label(aes(x = x, y = y, label = n, hjust = h),
              vjust = 1, data = ct, inherit.aes = F) +
    scale_color_manual(values = c('orange','forestgreen', 'red',  'black')) +
    labs(x = x, y = ylab, title = ttl) + theme(axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.title = element_blank())
}
```

# Results

Applying a modified DESeq2 approach, differential loop analysis was performed, assessing NTC versus Knockdown samples (M1 and M4 were pooled together). Differential loops from all chromosomes were also pooled together.

```{r,results_table,echo = FALSE,warning=FALSE,message=FALSE,out.width='100%'}
df_list %>% reactable()
```

# MA plot {.tabset}

## No cutoff
```{r,MA_plot_1,echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
ggplot(data=df_list,aes(x=baseMean,y=log2FoldChange)) +
  geom_point() +
  theme_minimal() +
  labs(y="log2FoldChange (NTC/Knockdown)",x="Normalized Mean Counts") +
  geom_hline(yintercept=0,color = "red")
```

## Cutoff = 400
Using a cutoff of 400 normalized mean counts
```{r,MA_plot_2,echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
ggplot(data=df_list[df_list$baseMean > 400,],aes(x=baseMean,y=log2FoldChange)) +
  geom_point() +
  theme_minimal() +
  labs(y="log2FoldChange (NTC/Knockdown)",x="Normalized Mean Counts") +
  geom_hline(yintercept=0,color = "red")
```

## Cutoff = 1000
Using a cutoff of 100 normalized mean counts
```{r,MA_plot_3,echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
ggplot(data=df_list[df_list$baseMean > 1000,],aes(x=baseMean,y=log2FoldChange)) +
  geom_point() +
  theme_minimal() +
  labs(y="log2FoldChange (NTC/Knockdown)",x="Normalized Mean Counts") +
  geom_hline(yintercept=0,color = "red")
```

## Cutoff = 1500
Using a cutoff of 100 normalized mean counts
```{r,MA_plot_4,echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
ggplot(data=df_list[df_list$baseMean > 1500,],aes(x=baseMean,y=log2FoldChange)) +
  geom_point() +
  theme_minimal() +
  labs(y="log2FoldChange (NTC/Knockdown)",x="Normalized Mean Counts") +
  geom_hline(yintercept=0,color = "red")
```

# Volcano plot

```{r,volcano_plot,echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
volc(df_list,'log2FoldChange','padj','-log10(FDR)','')
```
