---
title: ""
author: ""
date: ""
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 3
    smooth_scroll: false
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(plotly)
library(ggplot2)
library(data.table)
library(DESeq2)
library(reactable)
setwd("~/Documents/MED4_collab")
```

```{r,loading_R_objects,include = FALSE}
volc <- function(r, x, y, ylab, ttl) {
  d <- as.data.frame(r) %>%
    dplyr::rename(x = !!x, y = !!y) %>%
    mutate(kind = case_when((abs(x) > 1 & y < .05) ~ 'DE',
                            abs(x) > 1 ~ '|logFC|>1',
                            y < .05 ~ 'FDR<0.05',
                            T ~ 'NS'),
           y = -log10(y))
  ct <- d %>% 
    dplyr::filter(kind == 'DE') %>%
    mutate(up = x > 0) %>%
    dplyr::count(up) %>%
    mutate(x = ifelse(up, Inf, -Inf),
           y = Inf,
           h = as.numeric(up))
  ggplot(d, aes(x, y, color = kind)) +
    geom_point(alpha = .5) +
    geom_label(aes(x = x, y = y, label = n, hjust = h),
               vjust = 1, data = ct, inherit.aes = F) +
    scale_color_manual(values = c('orange','forestgreen', 'red',  'black')) +
    labs(x = "log2FoldChange (MED4KD/NTC)", y = ylab, title = ttl) + theme(axis.line = element_line(colour = "black"),
                                               panel.grid.major = element_blank(),
                                               panel.grid.minor = element_blank(),
                                               panel.border = element_blank(),
                                               panel.background = element_blank(),
                                               legend.title = element_blank())
}
#
mdat <- data.frame(
  sample = c("NTC-1","NTC-2","NTC-3","M1_1","M1-2","M1-3","M4-1","M4-2","M4-3"),
  condition = c("NTC","NTC","NTC","KD","KD","KD","KD","KD","KD")
) %>%
  column_to_rownames("sample")
# save
load(file="~/Documents/MED4_collab/data/HiCDCPlus/binsize10000/mustache/dds.RData")
load(file="~/Documents/MED4_collab/data/HiCDCPlus/binsize10000/mustache/vst.RData")
# load
# load(file="~/Documents/MED4_collab/data/HiCDCPlus/binsize10000/dds.RData")
# load(file="~/Documents/MED4_collab/data/HiCDCPlus/binsize10000/vst.RData")
# load(file="~/Documents/MED4_collab/data/HiCDCPlus/binsize10000/resLFC.RData")
```

```{r,preprocess,include=FALSE}
# dds <- rbind(mustache_binsize10000_chr1_DESeq2_obj,mustache_binsize10000_chr2_DESeq2_obj,mustache_binsize10000_chr3_DESeq2_obj,mustache_binsize10000_chr4_DESeq2_obj,mustache_binsize10000_chr5_DESeq2_obj,mustache_binsize10000_chr6_DESeq2_obj,mustache_binsize10000_chr7_DESeq2_obj,mustache_binsize10000_chr8_DESeq2_obj,mustache_binsize10000_chr9_DESeq2_obj,mustache_binsize10000_chr10_DESeq2_obj,mustache_binsize10000_chr11_DESeq2_obj,mustache_binsize10000_chr12_DESeq2_obj,mustache_binsize10000_chr13_DESeq2_obj,mustache_binsize10000_chr14_DESeq2_obj,mustache_binsize10000_chr15_DESeq2_obj,mustache_binsize10000_chr16_DESeq2_obj,mustache_binsize10000_chr17_DESeq2_obj,mustache_binsize10000_chr18_DESeq2_obj,mustache_binsize10000_chr19_DESeq2_obj)
#
# vst <- varianceStabilizingTransformation(dds)
#
# rd <- dds %>%
#   results(contrast = c('condition', 'KD', 'NTC')) %>%
#   data.frame() %>%
#   na.omit()
#
# s <- list.files(path = "~/Documents/MED4_collab/data/HiCDCPlus/binsize10000/mustache", pattern = '.txt',full.names = FALSE,recursive = FALSE) %>%
#   tibble(f = .) %>%
#   separate(f,c(NA,NA,NA,NA,"chr"),'\\_',F) %>%
#   mutate(f = file.path('~/Documents/MED4_collab/data/HiCDCPlus/binsize10000/mustache', f))
# s$chr <- gsub(".txt","",s$chr)
# d <- deframe(s[,c('chr', 'f')])
# rd_list <- lapply(d,fread) %>%
#   rbindlist()
# write.csv(rd_list,file="~/Documents/MED4_collab/data/HiCDCPlus/binsize10000/mustache/rd_list.csv",row.names = FALSE)
# library(pheatmap)
# vst_mat <- assay(vst)
# vst_cor <- cor(vst_mat)
# pheatmap(vst_cor)
# #
# rv <- rowVars(assay(vst))
# select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
# vst_mat <- assay(vst)[select,]
# vst_cor <- cor(vst_mat)
# pheatmap(vst_cor)
```

# Results

Loops were firstly called using [Mustache](https://genomebiology.biomedcentral.com/articles/10.1186/s13059-020-02167-0). Afterwards, a union of significantly called loops in each condition (M1, M4 and NTC) was used for differential analysis using a modified DESeq2 approach via [HiCDCPlus](https://www.nature.com/articles/s41467-021-23749-x). Knockdown samples (M1 and M4) were pooled together. Loops can be visualized using Hi-C contact maps found [here](http://206.12.101.70:8888/l/?d=IX1b0c4jQCyMPkO_xCqraQ).

```{r echo=FALSE,fig.cap="The number of significant loops (FDR < 0.05) is reported in each condition."}
xfun::embed_file("~/Documents/MED4_collab/data/HiCDCPlus/binsize10000/mustache/diff_final.csv")
dt <- data.frame("M1" = 9028,
           "M4" = 9089,
           "NTC"= 8752)
rownames(dt) <- "#_of_sig_loops"
reactable(dt)
# diff <- read.csv("~/Documents/MED4_collab/data/HiCDCPlus/binsize10000/mustache/rd_list.csv")
# diff$index <- sprintf("%s:%d:%d",diff$chr,diff$startI,diff$startJ)
# # df$coordinates1 <- sprintf("%s:%d-%d",df$BIN1_CHR,df$BIN1_START,df$BIN1_END)
# loops <- read.csv(file="~/Documents/MED4_collab/data/HiCDCPlus/binsize10000/mustache/mustache_loops.csv")
# loops$index <- sprintf("%s:%d:%d",loops$BIN1_CHR,loops$BIN1_START,loops$BIN2_START)
# loops_final <- loops %>%
#   dplyr::select(c("coordinates1","coordinates2","index"))
# diff_final <- left_join(diff,loops_final,by="index") %>%
#   dplyr::distinct()
# write.csv(diff_final,file="~/Documents/MED4_collab/data/HiCDCPlus/binsize10000/mustache/diff_final.csv",row.names = FALSE)
```

```{r echo=FALSE,include=FALSE}
# xfun::embed_file(file="~/Documents/MED4_collab/data/HiCDCPlus/binsize10000/mustache/mustache_loops.csv")
```

# PCA {.tabset}

## All loops
```{r,PCA,echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rv <- rowVars(assay(vst))
pc <- prcomp(t(assay(vst)))
condition <- mdat$condition
scores <- data.frame(pc$x, condition)
percentage <- round(pc$sdev / sum(pc$sdev) * 100, 2)
percentage <- paste( colnames(scores), "(", paste( as.character(percentage), "%", ")", sep="") )
p <- plot_ly(scores,x=scores$PC1,y=scores$PC2,text=rownames(scores),mode="markers",color=factor(condition),marker=list(size=11),colors = "Dark2")
p <- layout(p,title="",   xaxis = list(title = percentage[1]),
            yaxis = list(title = percentage[2]))
p
```

## Top 500 most variable loops
```{r,PCA_500,echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
rv <- rowVars(assay(vst))
select <- order(rv, decreasing=T)[seq_len(min(500,length(rv)))]
pc <- prcomp(t(assay(vst)[select,]))
condition <- mdat$condition
scores <- data.frame(pc$x, condition)
percentage <- round(pc$sdev / sum(pc$sdev) * 100, 2)
percentage <- paste( colnames(scores), "(", paste( as.character(percentage), "%", ")", sep="") )
p <- plot_ly(scores,x=scores$PC1,y=scores$PC2,text=rownames(scores),mode="markers",color=factor(condition),marker=list(size=11),colors = "Dark2")
p <- layout(p,title="",   xaxis = list(title = percentage[1]),
            yaxis = list(title = percentage[2]))
p
```

# MA plot {.tabset}

Logfold change is <b> MED4KD over NTC </b>

## No cutoff
```{r,MA_plot_1,echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
# resLFC <- DESeq2::lfcShrink(dds=dds, contrast=c("condition","NTC","KD"),type = "ashr")
DESeq2::plotMA(dds, ylim=c(-2,2),colSig="blue",xlab="Mean of Normalized Counts",alpha=0.05)
```

## Cutoff = 50
Using a cutoff of 50 normalized mean counts
```{r,MA_plot_2,echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
res <- results(dds,contrast=c("condition","KD","NTC")) %>%
  as.data.frame() %>%
  na.omit() %>%
  rownames_to_column("loops") %>%
  dplyr::select(c("baseMean","log2FoldChange","padj"))
res$baseMean <- as.numeric(res$baseMean)
res$log2FoldChange <- as.numeric(res$log2FoldChange)
res$significant <- res$padj < 0.05
res$significant <- as.logical(res$significant)
res <- res %>%
  dplyr::select(c("baseMean","log2FoldChange","significant"))
DESeq2::plotMA(res[res$baseMean > 50,], ylim=c(-2,2),colSig="blue",xlab="Mean of Normalized Counts",alpha=0.05)
```

## Cutoff = 100
Using a cutoff of 100 normalized mean counts
```{r,MA_plot_3,echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
DESeq2::plotMA(res[res$baseMean > 100,], ylim=c(-2,2),colSig="blue",xlab="Mean of Normalized Counts",alpha=0.05)
```

# Volcano plot

```{r,volcano_plot,echo = FALSE, warning=FALSE,message=FALSE,out.width='100%'}
res <- results(dds,contrast=c("condition","KD","NTC")) %>%
  as.data.frame() %>%
  na.omit() %>%
  rownames_to_column("loops")
volc(res,'log2FoldChange','padj','-log10(FDR)','')
```
